Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-27 ms.tex
more

Kinetic feedback occurs from an outflow of mass or energy  from the viscinity of the black hole in the form of a disk wind or jet, which deposits its energy across a much wider range of radii.  If the total outflow power is proportional to the SMBH accretion rate, $L_{\rm j} = \epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an efficiency factor.  Assuming that this energy is deposited uniformly out to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at the stagnation radius given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\end{eqnarray} where we have used the fact that $\dot{M} \approx \eta
M_{\rm enc}(r_{\rm s})/t_{\rm h}$ and that $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ for radii interior to the outer
break radius (eq.~[\ref{eq:rhostar}]).  Thus for $r_{\rm heat} \sim
r_{\rm s}$, even a modest heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other
sources of non-accretion powered heating.  However, if the jet distributes its energy on much larger scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ kpc $\gtrsim 10^{3}r_{\rm s}$ then the heating rate would be much less significant, even for an efficient jet ($\epsilon_{\rm j} \sim 1$).

@@

The total heating rate $\vwO = \sqrt{v_{w}^{\star \, 2} + v_{\rm
    w,0}^{\rm Ia \, 2} + v_{w}^{\rm p \, 2} + v_{\rm
    w,0}^{\bullet \, 2}}$ has
contributions from stellar winds, supernovae, pulsars, and SMBH
feedback. %  Since pulsar heating appears unlikely to contribute
% appreciably, and since SMBH heating is uncertain, we first focus on
% heating from stellar winds and supernovae.

Each of these heating sources will depend on the integrated star
formation history. This in term will will vary systematically with
galaxy mass and thus with $\Mbh$.  
%Qualitatively smaller galaxies willhave recent star formation and
% energetic winds from Wolf-Rayet stars will dominate the mass
% budget--$\vwO\gsim 1000 $ km/s.  On the other hand for larger galaxies
% will have older populations, and heating will be dominated by Ia
% supernovae on large scales and SMBH feedback on small scales.
We calculate the star formation history as a function of $\Mbh$ and
the $\vwO$ for each heating source using the cosmic star formation
histories in \citet{MosterNaab+:2013a}.  For $v_{w}^{\rm p}$ we
use equation~\eqref{eq:vmsp}. The star formation history simply enters
in through $\eta$. The effective $\eta$ for a given star formation
history will be the convolution of the star formation history with
$\eta_{\rm imp}\simeq $ {\bf AG: should have expression for $\eta_{\rm
    Imp}$ in appendix. Have to be careful as we also have the
  simplified expression for $\eta_{\rm Imp}$} ($\eta$ in the impulsive
limit). For $v_{w}^{\rm Ia}$ we use
equation~\eqref{eq:vIa}. Besides $\eta$ there is additional dependence
on the star formation history from the Ia rate--which is a convolution
of the star formation history with the delay-time distribution from
\citet{MaozMannucci+:2012a}. The calculation of $v_{w}^{\star}$
is somewhat involved: we leave the detailed description for
Appendix~\ref{app:windheat}. {\bf AG: still need a description of the
  Compton heating. In general, I struggled with the level of detail
  here. I couldn't think of a natural place to put the details for the
  other heating sources}

$\vwO$ from each heating source (stars, MSPs, SNe Ia, and black hole
feedback) is plotted in as a function of $\Mbh$ in
Figure~\ref{fig:vwSources}. $v_{w}^{\rm p}$ is negligible across
all $\Mbh$. For $\Mbh\lsim 10^{7.5} \Msun$, have substantial recent
star formation and $v_{w}^{\star}$ is dominant. For $\Mbh\gsim
10^{7.5} \Msun$ , the younger stellar population with their energetic
stellar winds is dead, and Compton heating plays an important role.

For reference we also plot the inferred $\eta$ for each $\Mbh$ in
Figure~\ref{fig:eta}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{eta.pdf}
\caption{\label{fig:eta} $\eta$ vs. $\Mbh$.}
\end{figure}

>>>>

Kinetic feedback occurs from an outflow of mass or energy  from the viscinity of the black hole in the form of a disk wind or jet, which deposits its energy across a much wider range of radii.  If the total outflow power is proportional to the SMBH accretion rate, $L_{\rm j} = \epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an efficiency factor.  Assuming that this energy is deposited uniformly out to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at the stagnation radius given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} where we have used the fact that $\dot{M} \approx \eta
M_{\rm enc}(r_{\rm s})/t_{\rm h}$ and that $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ for radii interior to the outer
break radius (eq.~[\ref{eq:rhostar}]).  Thus for $r_{\rm heat} \sim
r_{\rm s}$, even a modest heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other
sources of non-accretion powered heating.  However, if the jet distributes its energy on much larger scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ kpc $\gtrsim 10^{3}r_{\rm s}$ then the heating rate would be much less significant, even for an efficient jet ($\epsilon_{\rm j} \sim 1$).

@@

The total gas heating rate $\vwO$ includes contributions from stellar winds, supernovae, pulsars, and SMBH feedback.  The strength of each heating source will depend on the SMBH mass and the star formation history in the galactic nuclear region, the latter of which will vary systematically with the galaxy mass and hence with $\Mbh$.  Qualitatively, smaller galaxies will on average have more recent star formation, resulting in energetic young stellar winds and supernovae which dominate the energy budget, resulting in a high heating rate of $\vwO \sim 1000 $ km s$^{-1}$.  Larger galaxies, on the other hand, will on average have older stellar populations, with their heating rates dominated by Ia supernovae (on large radial scales) and SMBH feedback.

The average star formation history as a function of $\Mbh$ and the $\vwO$ for each heating source is estimated by using the average cosmic star formation histories determined by \citet{MosterNaab+:2013a}.  As described in detail in Appendix \ref{app:windheat}, the average star formation history is used to estimate the stellar wind heating $v_{\rm w}^{\star}$ and the mass input parameter $\eta$ (Fig.~\ref{fig:eta}), the latter of which affects the other sources of feedback, including AGN ($v_{\rm w}^{\bullet}$; eqs.~[\ref{eq:vC}], [\ref{eq:vjet}]), millisecond pulsars ($v_{w}^{\rm msp}$; eq.~[\ref{eq:vmsp}]), and Ia supernovae ($v_{w}^{\rm Ia}$; eq.~[\ref{eq:vIa}]), the latter of which also depends on a convolution of the star formation history and the Ia delay-time distribution (eq.~[\ref{eq:DTD}])

Figure~\ref{fig:vwSources} shows $\vwO$ from each heating source (stars, MSPs, SNe Ia, and black hole
feedback) as a function of black hole mass.  MSP heating is found to be negligible at all black hole masses.  The young stellar populations in galaxies with $\Mbh\lsim 10^{7.5} \Msun$ results in $v_{w}^{\star}$ being dominant.  For $\Mbh\gsim 10^{7.5} \Msun$ , the younger stellar population with their energetic
stellar winds is dead, and Compton heating plays an important role.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{eta.pdf}
\caption{\label{fig:eta} Parameter characterizing stellar mass loss $\eta$ (eq.~[\ref{eq:q}]) as a function of black hole mass $\Mbh$, calculated for average star formation histories corresponding to galaxies with halo masses corresponding to $\Mbh$ according to the results of \citet{MosterNaab+:2013a}. }
\end{figure}
